Analytical Predictions of Exoplanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-19 writeups/Su_disk_disp.tex
stash

\subsubsection{Governing Principles: Evolution of Enclosed Phase Space
Area}\label{sss:a_evo}

First, we consider how the enclosed phase space area by a trajectory evolves
over time. In the absence of separatrix encounters, the enclosed phase space
area $\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. In
particular, we adopt convention where
\begin{equation}
    A \equiv \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.\label{eq:a_oint}
\end{equation}
This definition of $A$ has the advantage of (i) being continuous at $\cos \theta
= 1$, and (ii) being easily expressible as combinations of the $A_{\rm i}$ of
Eqs.~\eqref{se:area_ward}. The bounds of the integral are either a libration
($\phi$ returns to its original value with the same $\dot{q\phi}$ sign) or
circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly that $A$ has
simple physical interpretation of are enclosed on the unit sphere by $\uv{s}$
over one cycle measured relative to $+\uv{z}$:
\begin{equation}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation}

Observe that when $\eta_{\rm i} \gg 1$, trajectories librate about $\uv{l}_{\rm d}$
with constant $\theta_{\rm sd}$, meaning they enclose initial phase space area
\begin{equation}
    A_{\rm i} \equiv \at{A}_{t = 0}
        = 2\pi\p{1 - \cos \theta_{\rm sd, i}}.\label{eq:ai_qsd}
\end{equation}
Complications arise when considering finite $\eta_{\rm i}$, as trajectories near CS2
or CS3 librate about these equilibria respectively, rather than $\uv{l}_{\rm
d}$, and Eq.~\eqref{eq:ai_qsd} is no longer exact. In practice,
Eq.~\eqref{eq:ai_qsd} holds very well when defining $\theta_{\rm sd, i}$ as the
angular distance to CS2; an exception is discussed in Section~\ref{sss:evol_traj}.

>>>>

\subsubsection{Governing Principle: Evolution of Enclosed Phase Space
Area}\label{sss:a_evo}

First, we consider how the enclosed phase space area by a trajectory evolves
over time. In the absence of separatrix encounters, the enclosed phase space
area $\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. We adopt
convention where
\begin{equation}
    A \equiv \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.\label{eq:a_oint}
\end{equation}
This definition of $A$ has the advantage of (i) being continuous across
transitions from circulating to librating that cross the North pole
($\cos \theta = 1$), and (ii) being easily expressible as combinations of the
$A_{\rm i}$ of Eqs.~\eqref{se:area_ward}. The bounds of the integral are either
a libration or circulation cycle. Note that when $\eta_{\rm i} \gg 1$,
trajectories librate about $\uv{l}_{\rm d}$ with constant $\theta_{\rm sd}$,
meaning they enclose initial phase space area
\begin{equation}
    A_{\rm i} = 2\pi\p{1 - \cos \theta_{\rm sd, i}}.\label{eq:ai_qsd}
\end{equation}
Complications arise when considering finite $\eta_{\rm i}$, as trajectories near
CS2 or CS3 librate about these equilibria, rather than $\uv{l}_{\rm d}$, and
Eq.~\eqref{eq:ai_qsd} is no longer exact. In practice, Eq.~\eqref{eq:ai_qsd}
holds very well when defining $\theta_{\rm sd, i}$ as the angular distance to
CS2; an exception is discussed in Section~\ref{sss:evol_traj}.
